---
title: "Amott_Hw1"
output: 
  html_document: 
    keep_md: yes
---
```{r}
#install.packages("tidyverse")
```
```{r}
library("tidyverse")
```
1. Navigate to the R console and calculate the following expressions.  
  + 5 - 3 * 2  
  + 8 / 2 ** 2  
  
```{r}
5 - 3 * 2
8 / 2 **2

```
  
2. Did any of the results in #1 surprise you? Write two programs that calculate each expression such that the result for the first example is 4 and the second example is 16.  
```{r}
(5 - 3) * 2
(8 / 2)**2
```

3. Make a new object `pi` as 3.14159265359. 
```{r}
pi <- 3.14159265359
```

4. Is `pi` an integer or numeric? Why? Show your code.  
```{r}
class(pi)
```
5. You have decided to use your new analytical powers in R to become a professional gambler. Here are your winnings and losses this week. Note that you don't gamble on the weekends!  
```{r, eval=F, echo=T}
blackjack <- c(140, -20, 70, -120, 240, NA, NA)
roulette <- c(60, 50, 120, -300, 10, NA, NA)
```

a. Build a new vector called `days` for the days of the week.
ˆ
```{r}
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

```{r, eval=F, echo=T}
names(blackjack) <- days
names(roulette) <- days
```

b. Calculate how much you won or lost in blackjack over the week. 
```{r}
sum(blackjack)
```
c. The result was NA because there is missing data for saturday and sunday. We can recalculate it with the NA removed.

```{r}
sum(blackjack, na.rm=TRUE)
```

d. Calculate how much you won or lost in roulette over the week. 
```{r}
sum(roulette, na.rm = TRUE)
```

e. Build a `total_week` vector to show how much you lost or won on each day over the week. Which days seem lucky or unlucky for you?
```{r}
total_week <- c(days)
days[1]
sum(blackjack[1]+roulette[1])
days[2]
sum(blackjack[2]+roulette[2])
days[3]
sum(blackjack[3]+roulette[3])
days[4]
sum(blackjack[4]+roulette[4])
days[5]
sum(blackjack[5]+roulette[5])
```
Monday, Wednesday, Friday are lucky days and thursday is very unlucky.

f. Should you stick to blackjack or roulette? Write a program that verifies this below.  
```{r}
sum(blackjack, na.rm = TRUE)
sum(roulette, na.rm = TRUE)
sum(blackjack, na.rm = TRUE) > sum(roulette, na.rm = TRUE)

```
Stick to Blackjack!

